### 2019-03-11 Thesis_Latex/Thesis/Chapter4/Chapter4.tex
update thesis

\chapter{Moving Obstacle Avoidance}

\section{Problem Formulation}
The collision avoidance problem is very dependent on the vehicle modeling since it is a requirement for  adaptive MPC law design. Figure \ref{fig:obstacleAvoidance} illustrates a typical scenario of overtaking of a moving obstacle.
% COLLISION AVOIDANCE FIGURE
\begin{figure}[!h]
	\centering
	\includegraphics[width=\textwidth]{../figure/obstacleAvoidance/obstacleAvoidance.pdf}
	\caption{Problem description of collision avoidance on a road with only two lanes.}
	\label{fig:obstacleAvoidance}
\end{figure}

The model used in this report should take into account the kinematic and dynamic aspects of the vehicle. Here, we present a non linear mathematical model of a vehicle used for the development of a collision avoidance system.
The model has four states and two inputs:
\begin{equation}
\begin{array}{cc}
\vec{x}=\begin{bmatrix}
x\\y\\\theta\\v 
\end{bmatrix},\qquad 
\vec{u}=\begin{bmatrix}
T\\\delta 
\end{bmatrix}
\end{array} 
\end{equation}
where $(x,y)$ are the global coordinates of the contact point between the rear wheel and the ground, $\theta$ is the heading angle of the car body with respect to the $x$-axis and $v$ is the linear speed of the car (positive). The manipulated variables are $T$ the throttle (positive when accelerating/negative when braking) and $\delta$ the steering angle of the front wheel with respect to the vehicle ($0$ when aligned with car, counterclockwise positive). Figure \ref{fig:car_model} illustrates the applied nonlinear bicycle model and the related parameters.
% BICYCLE MODEL
\begin{figure}[!h]
	\centering
	\includegraphics[width=0.80\textwidth]{../figure/car_model.pdf}
	\caption{Bicycle model of a car (adapted from \cite{siciliano}).}
	\label{fig:car_model}
\end{figure}

The ATLASCAR2 can be modeled using the non-linear kinematic bicycle model described by the following equations of motion \cite{safety} \cite{swarms}:
\begin{equation}
\label{eqn:dynamics_model_obstacle_avoidance}
\left \{ \begin{array}{llll}
\dot{x} = v\cos(\theta)\\
\dot{y} = v\sin(\theta)\\
\dot{\theta} =\dfrac{v}{C_L}\tan(\delta)\\
\dot{v} =0.5 \cdot T
\end{array} 
\right .
\Longrightarrow 
\begin{array}{llll}
\dot{\vec{x}} = \vec{f}(\vec{x},\vec{u})\\
\vec{y} = \vec{g}(\vec{x},\vec{u})
\end{array}
\end{equation}
where $C_L$ is the car length. In order to simplify the model, it is assumed that only the front wheel can be steered. Moreover, in this paper it is assumed that the ATLASCAR2 does not slip, so any slippage is thus considered as an external disturbance. Under this assumption, the slip angle is zero, meaning that the velocity is directed along the heading of the vehicle. In order to use MPC, the state space model needs to be linearized with a first order approximation and also re-written in a more compact form:
\begin{equation}
\label{eqn:dynamics_model_non_linear}
\begin{array}{llll}
\dot{\vec{x}} = \vec{f}(\vec{x},\vec{u})\\
\vec{y} = \vec{g}(\vec{x},\vec{u})
\end{array} \Longrightarrow
\begin{array}{ll}
\dot{\vec{x}} =\vec{A}_c \vec{x}+ \vec{B}_c \vec{u}\\
\vec{y} =\vec{C}_c \vec{x} + \vec{D}_c \vec{u}
\end{array}
\end{equation}
where the matrices $\vec{A}_c$, $\vec{B}_c$, $\vec{C}_c$ and $\vec{D}_c$ are obtained as follows:
\begin{equation}
\begin{array}{ccc}
\vec{A}_c=\displaystyle\frac{\partial \vec{f}(\vec{x},\vec{u})}{\partial \vec{x}}=\begin{bmatrix}
0&0&-v\sin(\theta)&\cos(\theta)\\
0&0&v\cos(\theta)&\sin(\theta)\\
0&0&0&\tan(\delta)/C_L\\
0&0&0&0
\end{bmatrix},
\\\\
\vec{B}_c=\displaystyle\frac{\partial \vec{f}(\vec{x},\vec{u})}{\partial \vec{u}}=\begin{bmatrix}
0&0\\
0&0\\
0&\dfrac{v}{C_L}\big(\tan(\delta)^2+1\big)\\
0.5&0
\end{bmatrix},
\\\\
\vec{C}_c=\displaystyle\frac{\partial \vec{g}(\vec{x},\vec{u})}{\partial \vec{u}} = \mathbf{I}_4, 
\qquad
\vec{D}_c=\frac{\partial \vec{g}(\vec{x},\vec{u})}{\partial \vec{u}}=\mathbf{0}_{4\times2}.
\end{array}
\end{equation} 
The simple linearized approximation of the system to describe the dynamics of the ATLASCAR2 will be evaluated at the operating conditions. Note also that the system we are considering is a linear state-space model whose dynamics vary as a function of certain time-varying parameters. The system to be controlled is usually modeled by a discrete state-space model in the MPC literature. Therefore, (\ref{eqn:dynamics_model_non_linear})
is transformed into a discrete state-space model to be used by the Model Predictive Controller:
\begin{equation}
\label{eqn:dynamics_ss_obstacle_avoidance_dis}
\begin{array}{ll}
\dot{\vec{x}} =\vec{A}_c \vec{x}+ \vec{B}_c \vec{u}\\
\vec{y} =\vec{C}_c \vec{x} + \vec{D}_c \vec{u}
\end{array}
\Longrightarrow
\begin{array}{rr}
{\vec{x}}(k+1) =\vec{A}_d \vec{x}(k)+ \vec{B}_d \vec{u}(k)\\
\vec{y}(k) =\vec{C}_d \vec{x}(k) + \vec{D}_d \vec{u}(k)
\end{array}
\end{equation}

>>>>

\chapter{Moving Obstacle Avoidance}
In this chapter we present an Obstacle Avoidance alghorithm based on Adaptive Model Predictive Control. First we introduced the case-study model, then we designed the controller and the method of decision-making. Finally we verified the proposed strategy in different scenarios.

\section{Problem Formulation}
The collision avoidance problem is very dependent on the vehicle modeling since it is a requirement for  adaptive MPC law design. Figure \ref{fig:obstacleAvoidance} illustrates a typical scenario of overtaking of a moving obstacle. The maneuver from the driver is to temporarily move to another lane, drive past the obstacle, and move back to the original lane afterward. 
% COLLISION AVOIDANCE FIGURE
\begin{figure}[!h]
	\centering
	\includegraphics[width=\textwidth]{../figure/obstacleAvoidance/obstacleAvoidance.pdf}
	\caption{Problem description of collision avoidance on a road with only two lanes.}
	\label{fig:obstacleAvoidance}
\end{figure}

This is possible if and only if there is a free lane or  sufficient space for overtaking the obstacle. In the case the road is busy with vehicles, the ATLASCAR2 must slow down and adapts its speed to that of the closest obstacle until the scenario changes.

\subsection{Car-like Model}
The model used in this part of the thesis should take into account the kinematic and dynamic aspects of the vehicle. Here, we present a non linear mathematical model of a vehicle used for the development of a collision avoidance system.

The model has four states and two inputs:
\begin{equation}
\begin{array}{cc}
\vec{x}=\begin{bmatrix}
x\\y\\\theta\\v 
\end{bmatrix},\qquad 
\vec{u}=\begin{bmatrix}
T\\\delta 
\end{bmatrix}
\end{array} 
\end{equation}
where $(x,y)$ are the global coordinates of the contact point between the rear wheel and the ground, $\theta$ is the heading angle of the car body with respect to the $x$-axis and $v$ is the linear speed of the car (positive). The manipulated variables are $T$ the throttle (positive when accelerating/negative when braking) and $\delta$ the steering angle of the front wheel with respect to the vehicle ($0$ when aligned with car, counterclockwise positive). Figure \ref{fig:car_model} illustrates the applied nonlinear bicycle model and the related parameters.
% BICYCLE MODEL
\begin{figure}[!h]
	\centering
	\includegraphics[width=0.80\textwidth]{../figure/car_model.pdf}
	\caption{Bicycle model of a car (adapted from \cite{siciliano}).}
	\label{fig:car_model}
\end{figure}

The ATLASCAR2 can be modeled using the non-linear kinematic bicycle model described by the following equations of motion \cite{safety} \cite{swarms}:
\begin{equation}
\label{eqn:dynamics_model_obstacle_avoidance}
\left \{ \begin{array}{llll}
\dot{x} = v\cos(\theta)\\
\dot{y} = v\sin(\theta)\\
\dot{\theta} =\dfrac{v}{C_L}\tan(\delta)\\
\dot{v} =0.5 \cdot T
\end{array} 
\right .
\Longrightarrow 
\begin{array}{llll}
\dot{\vec{x}} = \vec{f}(\vec{x},\vec{u})\\
\vec{y} = \vec{g}(\vec{x},\vec{u})
\end{array}
\end{equation}

where $C_L$ is the car length. When the velocity is zero then the rate of change of heading angle is zero, that is, it is not possible to change the vehicle's orientation
when it is not moving. If the front wheel is orthogonal to the back wheel, i.e. the steering angle is $\frac{\pi}{2}$, the ATLASCAR2 cannot move forward and the model enters an undefined region. In order to simplify the model, it is assumed that only the front wheel can be steered. Moreover, in this paper it is assumed that the ATLASCAR2 does not slip, so any slippage is thus considered as an external disturbance. Under this assumption, the slip angle is zero, meaning that the velocity is directed along the heading of the vehicle. In particular the rate of change of heading angle is referred to as turn rate which can be evaluated with a gyroscope. This yaw rate can also be deduced from the angular velocity of the wheels on the left- and right-hand sides of the vehicle which therefore rotate at different speeds because follow arcs of different radius. We can write the non-holonomic constraint which is an expression for velocity in the vehicle's $y$-direction in the world coordinate frame as follows:
\begin{equation}
	\dot{y}\cos\theta-\dot{x}\cos\theta=0
\end{equation}
This equation cannot be integrated to form a
relationship between $x,y$ and $\theta$.

In order to use an MPC controller, the state space model needs to be linearized with a first order approximation and also re-written in a more compact form:
\begin{equation}
\label{eqn:dynamics_model_non_linear}
\begin{array}{llll}
\dot{\vec{x}} = \vec{f}(\vec{x},\vec{u})\\
\vec{y} = \vec{g}(\vec{x},\vec{u})
\end{array} \Longrightarrow
\begin{array}{ll}
\dot{\vec{x}} =\vec{A}_c \vec{x}+ \vec{B}_c \vec{u}\\
\vec{y} =\vec{C}_c \vec{x} + \vec{D}_c \vec{u}
\end{array}
\end{equation}
where the matrices $\vec{A}_c$, $\vec{B}_c$, $\vec{C}_c$ and $\vec{D}_c$ are obtained as follows:
\begin{equation}
\begin{array}{ccc}
\vec{A}_c=\displaystyle\frac{\partial \vec{f}(\vec{x},\vec{u})}{\partial \vec{x}}=\begin{bmatrix}
0&0&-v\sin(\theta)&\cos(\theta)\\
0&0&v\cos(\theta)&\sin(\theta)\\
0&0&0&\tan(\delta)/C_L\\
0&0&0&0
\end{bmatrix},
\\\\
\vec{B}_c=\displaystyle\frac{\partial \vec{f}(\vec{x},\vec{u})}{\partial \vec{u}}=\begin{bmatrix}
0&0\\
0&0\\
0&\dfrac{v}{C_L}\big(\tan(\delta)^2+1\big)\\
0.5&0
\end{bmatrix},
\\\\
\vec{C}_c=\displaystyle\frac{\partial \vec{g}(\vec{x},\vec{u})}{\partial \vec{u}} = \mathbf{I}_4, 
\qquad
\vec{D}_c=\frac{\partial \vec{g}(\vec{x},\vec{u})}{\partial \vec{u}}=\mathbf{0}_{4\times2}.
\end{array}
\end{equation} 
The simple linearized approximation of the system to describe the dynamics of the ATLASCAR2 will be evaluated at the operating conditions. Note also that the system we are considering is a linear state-space model whose dynamics vary as a function of certain time-varying parameters. The system to be controlled is usually modeled by a discrete state-space model in the MPC literature. Therefore, (\ref{eqn:dynamics_model_non_linear})
is transformed into a discrete state-space model to be used by the Model Predictive Controller:
\begin{equation}
\label{eqn:dynamics_ss_obstacle_avoidance_dis}
\begin{array}{ll}
\dot{\vec{x}} =\vec{A}_c \vec{x}+ \vec{B}_c \vec{u}\\
\vec{y} =\vec{C}_c \vec{x} + \vec{D}_c \vec{u}
\end{array}
\Longrightarrow
\begin{array}{rr}
{\vec{x}}(k+1) =\vec{A}_d \vec{x}(k)+ \vec{B}_d \vec{u}(k)\\
\vec{y}(k) =\vec{C}_d \vec{x}(k) + \vec{D}_d \vec{u}(k)
\end{array}
\end{equation}

### 2019-03-11 Thesis_Latex/Thesis/Chapter5/Chapter5.tex
update thesis

Moreover the following figures show the trend of the main parameters confirming that the control strategy used allows the vehicle to follow the path. In particular we simulated also a small error in the sensor dynamics in order to make the simulation more realistic: we added a 3 percent error to the longitudinal velocity and this is evident from the small noise in the graphs of the steering angle (Figure \ref{fig:steering_laneFollowing}) and the lateral deviation (Figure \ref{fig:lateral_deviation_laneFollowing}).
Figure {\ref{fig:longitudinal_velocity_laneFollowing}} shows the evolution of the vehicle longitudinal velocity. At the start of the simulation, this velocity is equal to the initial condition for longitudinal velocity parameter $V_0$. At run time, we can note that $V_x$ reaches the predefined value of \SI{20}{m/s} meters for second and then it stabilizes near the cruising speed because it continues to vary the steering angle to adapt to the path to be followed. Finally we presents the overall scheme for the lane following developted in Simulink depicted in Figure \ref{fig:scheme_lane_following}.
%LANE FOLLOWING SINUSOIDAL PATH - SIGNALS
\begin{figure}[!h]
	\centering
	\begin{minipage}[t]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following/figure/LongitudinalVelocityVsTime.pdf}
		\subcaption{Longitudinal velocity $V_x$ w.r.t. time.}
		\label{fig:longitudinal_velocity_laneFollowing}
	\end{minipage}
	\begin{minipage}[t]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following/figure/AccelerationVsTime.pdf}
		\subcaption{Acceleration $u_1$ w.r.t. time.}
		\label{fig:acceleration_laneFollowing}
	\end{minipage}
	\begin{minipage}[t]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following/figure/SteeringAngleVsTime.pdf}
		\subcaption{Steering angle $u_2$ w.r.t. time.}
		\label{fig:steering_laneFollowing}
	\end{minipage}
	%\begin{minipage}[t]{\columnwidth}
	%	\includegraphics[width=\columnwidth]{../../MATLAB/lane_following/figure/RelativeYawAngleVsTime.pdf}
	%	\subcaption{Relative yaw angle $e_2$ w.r.t. time.}
	%	\label{fig:relative_yaw_angle_laneFollowing}
	%\end{minipage}
	\begin{minipage}[t]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following/figure/LateralDeviationVsTime.pdf}
		\subcaption{Lateral deviation $e_1$ w.r.t. time.}
		\label{fig:lateral_deviation_laneFollowing}
	\end{minipage}
	\caption{Time signals of the ATLASCAR2 in the simulation with a sinusoidal path.}
	\label{fig:laneFollowing_signals}
\end{figure}

@@

%LANE FOLLOWING SIMPLE CURVE - PATH/CURVATURE
\begin{figure}[!h]
	\centering
	\begin{minipage}[t]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following_curve/figure/LongitudinalVelocityVsTime_curve.pdf}
		\subcaption{Longitudinal velocity $V_x$ w.r.t. time.}
		\label{fig:longitudinal_velocity_laneFollowing_curve}
	\end{minipage}
	\begin{minipage}[t]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following_curve/figure/AccelerationVsTime_curve.pdf}
		\subcaption{Acceleration $u_1$ w.r.t. time.}
		\label{fig:acceleration_laneFollowing_curve}
	\end{minipage}
	\begin{minipage}[t]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following_curve/figure/SteeringAngleVsTime_curve.pdf}
		\subcaption{Steering angle $u_2$ w.r.t. time.}
		\label{fig:steering_laneFollowing_curve}
	\end{minipage}
	%\begin{minipage}[t]{\columnwidth}
	%	\includegraphics[width=\columnwidth]{../../MATLAB/lane_following_curve/figure/RelativeYawAngleVsTime_curve.pdf}
	%	\subcaption{Relative yaw angle $e_2$ w.r.t. time.}
	%	\label{fig:relative_yaw_angle_laneFollowing_curve}
	%\end{minipage}
	\begin{minipage}[t]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following_curve/figure/LateralDeviationVsTime_curve.pdf}
		\subcaption{Lateral deviation $e_1$ w.r.t. time.}
		\label{fig:lateral_deviation_laneFollowing_curve}
	\end{minipage}
	\caption{Time signals of the ATLASCAR2 in the simulation with a curve path.}
	\label{fig:laneFollowing_signals_curve}
\end{figure}

>>>>

Moreover the following figures show the trend of the main parameters confirming that the control strategy used allows the vehicle to follow the path. In particular we simulated also a small error in the sensor dynamics in order to make the simulation more realistic: we added a 3 percent error to the longitudinal velocity and this is evident from the small noise in the graphs of the steering angle (Figure \ref{fig:steering_laneFollowing}) and the lateral deviation (Figure \ref{fig:lateral_deviation_laneFollowing}).
Figure {\ref{fig:longitudinal_velocity_laneFollowing}} shows the evolution of the vehicle longitudinal velocity. At the start of the simulation, this velocity is equal to the initial condition for longitudinal velocity parameter $V_0$. At run time, we can note that $V_x$ reaches the predefined value of \SI{20}{m/s} and then it stabilizes near the cruising speed because it continues to vary the steering angle to adapt to the path to be followed. Finally we presents the overall scheme for the lane following developted in Simulink depicted in Figure \ref{fig:scheme_lane_following}.
%LANE FOLLOWING SINUSOIDAL PATH - SIGNALS
\begin{figure}[!h]
	\centering
	\begin{minipage}[t]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following/figure/LongitudinalVelocityVsTime.pdf}
		\subcaption{Longitudinal velocity $V_x$ w.r.t. time.}
		\label{fig:longitudinal_velocity_laneFollowing}
	\end{minipage}
	\begin{minipage}[t]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following/figure/AccelerationVsTime.pdf}
		\subcaption{Acceleration $u_1$ w.r.t. time.}
		\label{fig:acceleration_laneFollowing}
	\end{minipage}
	\begin{minipage}[t]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following/figure/SteeringAngleVsTime.pdf}
		\subcaption{Steering angle $u_2$ w.r.t. time.}
		\label{fig:steering_laneFollowing}
	\end{minipage}
	\begin{minipage}[t]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following/figure/RelativeYawAngleVsTime.pdf}
		\subcaption{Relative yaw angle $e_2$ w.r.t. time.}
		\label{fig:relative_yaw_angle_laneFollowing}
	\end{minipage}
	\begin{minipage}[t]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following/figure/LateralDeviationVsTime.pdf}
		\subcaption{Lateral deviation $e_1$ w.r.t. time.}
		\label{fig:lateral_deviation_laneFollowing}
	\end{minipage}
	\begin{minipage}[t]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following/figure/LateralVelocityVsTime.pdf}
		\subcaption{Lateral velocity $V_y$ w.r.t. time.}
		\label{fig:lateral_velocity_laneFollowing}
	\end{minipage}
	\caption{Time signals of the ATLASCAR2 in the simulation with a sinusoidal path.}
	\label{fig:laneFollowing_signals}
\end{figure}

@@

%LANE FOLLOWING SIMPLE CURVE - PATH/CURVATURE
\begin{figure}[!h]
	\centering
	\begin{minipage}[t]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following_curve/figure/LongitudinalVelocityVsTime_curve.pdf}
		\subcaption{Longitudinal velocity $V_x$ w.r.t. time.}
		\label{fig:longitudinal_velocity_laneFollowing_curve}
	\end{minipage}
	\begin{minipage}[t]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following_curve/figure/AccelerationVsTime_curve.pdf}
		\subcaption{Acceleration $u_1$ w.r.t. time.}
		\label{fig:acceleration_laneFollowing_curve}
	\end{minipage}
	\begin{minipage}[t]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following_curve/figure/SteeringAngleVsTime_curve.pdf}
		\subcaption{Steering angle $u_2$ w.r.t. time.}
		\label{fig:steering_laneFollowing_curve}
	\end{minipage}
	\begin{minipage}[t]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following_curve/figure/RelativeYawAngleVsTime_curve.pdf}
	\subcaption{Relative yaw angle $e_2$ w.r.t. time.}
		\label{fig:relative_yaw_angle_laneFollowing_curve}
	\end{minipage}
	\begin{minipage}[t]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following_curve/figure/LateralDeviationVsTime_curve.pdf}
		\subcaption{Lateral deviation $e_1$ w.r.t. time.}
		\label{fig:lateral_deviation_laneFollowing_curve}
	\end{minipage}
	\begin{minipage}[t]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following_curve/figure/LateralVelocityVsTime_curve.pdf}
		\subcaption{Lateral velocity $V_y$ w.r.t. time.}
		\label{fig:lateral_velocity_laneFollowing_curve}
	\end{minipage}
	\caption{Time signals of the ATLASCAR2 in the simulation with a curve path.}
	\label{fig:laneFollowing_signals_curve}
\end{figure}

### 2019-03-11 Thesis_Latex/Thesis/Introduction/Introduction.tex
update thesis

\section{Context of the Problem and Proposed Approach}\label{sec:context}
Dynamic environments pose several added difficulties to the motion planning problem. The dynamics of the ATLASCAR2 must be taken into account, and there are limitations due to the sensors range and uncertainty in measurements, that must be reflected on the motion plan. Besides it, a motion plan must incorporate time restrictions, meaning the vehicle will require a certain amount of time to accomplish a task. For example, when crossing a road, the ATLASCAR2 must do it fast enough to avoid incoming cars. The proposed algorithms were studied for the ATLASCAR2 project in which the group for Robotics and Automation at the University of Aveiro has setup and adapted a common commercial electric vehicle to provide a versatile  framework  to  develop  studies  and  research \cite{vsantos2010} \cite{vsantos2019}. The fact that the vehicle is electric allows to use the energy stored in the batteries, making it easier to power the sensors installed. In fact, the ATLACAR2 is equipped with sensors, such as lidar, that measure the distance to obstacles in front and around the vehicle. The obstacles can be static, such as a large pothole, or moving, such as a moving vehicle on the same or a nearby lane. The most common maneuver from the driver is to temporarily move to another lane, drive past the obstacle, and move back to the original lane afterward. In this case, we want to design an obstacle avoidance system that moves the ATLASCAR2 around a moving obstacle in the lane using throttle and steering angle. This system uses an adaptive Model Predictive Controller that updates both the predictive model and the mixed input/output constraints at each control interval. Moreover we want to develop a lane-keeping assist system for the vehicle: it has a sensor, such as camera or laser, that measures the lateral deviation and relative yaw angle between the centerline of a lane and the ATLASCAR2; it also measures the current lane curvature and its derivative. Depending on the curve length that the sensor can view, the curvature in front of the vehicle can be calculated from the current curvature and its derivative. This system keeps the autonomous car travelling along the centerline of the lanes on the road by adjusting the front steering angle. The goal for lane keeping control is to drive both lateral deviation and relative yaw angle close to zero.

>>>>

\section{Context of the Problem and Proposed Approach}\label{sec:context}
Dynamic environments pose several added difficulties to the motion planning problem. The dynamics of the ATLASCAR2 must be taken into account, and there are limitations due to the sensors range and uncertainty in measurements, that must be reflected on the motion plan. Besides it, a motion plan must incorporate time restrictions, meaning the vehicle will require a certain amount of time to accomplish a task. For example, when crossing a road, the ATLASCAR2 must do it fast enough to avoid incoming cars. The proposed algorithms were studied for the ATLASCAR2 project in which the group for Robotics and Automation at the University of Aveiro has setup and adapted a common commercial electric vehicle to provide a versatile  framework  to  develop  studies  and  research \cite{vsantos2010} \cite{vsantos2019}. The fact that the vehicle is electric allows to use the energy stored in the batteries, making it easier to power the sensors installed. In fact, the ATLACAR2 is equipped with sensors, such as lidar, that measure the distance to obstacles in front and around the vehicle. The obstacles can be static, such as a large pothole, or moving, such as a moving vehicle on the same or a nearby lane. The most common maneuver from the driver is to temporarily move to another lane, drive past the obstacle, and move back to the original lane afterward. In this case, we want to design an obstacle avoidance system that moves the ATLASCAR2 around  moving obstacles in the lane using throttle and steering angle. This system uses an adaptive Model Predictive Controller that updates both the predictive model and the mixed input/output constraints at each control interval. Moreover we want to develop a lane-keeping assist system for the vehicle: it has a sensor, such as camera or laser, that measures the lateral deviation and relative yaw angle between the centerline of a lane and the ATLASCAR2; it also measures the current lane curvature and its derivative. Depending on the curve length that the sensor can view, the curvature in front of the vehicle can be calculated from the current curvature and its derivative. This system keeps the autonomous car travelling along the centerline of the lanes on the road by adjusting the front steering angle. The goal for lane keeping control is to drive both lateral deviation and relative yaw angle close to zero.
